Design of Adaptive Constellations and Error Protection Coding for Wireless Network Coding in 5-node Butterfly Networks

### 2016-12-12 Const_design.tex
Renamed main body + added basic web presentation

%% LyX 2.1.5 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[letterpaper,english,10pt]{article}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{array}
\usepackage{float}
\usepackage{units}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[dvips]{graphicx}
\usepackage{setspace}
\usepackage{wasysym}
\usepackage{nomencl}
% the following is useful when we have the old nomencl.sty package
\providecommand{\printnomenclature}{\printglossary}
\providecommand{\makenomenclature}{\makeglossary}
\makenomenclature
\doublespacing

@@

\usepackage{babel}
\begin{document}

@@

\author{Pavel Prochazka, Tomas Uricar, David Halls and Jan Sykora}

\maketitle
{\footnotesize{}The corresponding author D. Halls is with the Toshiba
-- Telecommunication Research Laboratory (email: david.halls@toshiba-trel.com).
Pavel Prochazka, Tomas Uricar and Jan Sykora were with the Czech Technical
University in Prague, Czech Republic (e-mail: \textbackslash{}\{prochp10,
uricatom, Jan.Sykora\textbackslash{}\}@fel.cvut.cz). }{\footnotesize \par}

\thispagestyle{empty}

\pagebreak{}
\begin{abstract}
In this paper we focus on the design of a reliable communication scheme
for generic 5-node Wireless Butterfly Networks (WBN). Since the WBN
represents a natural extension of the basic 2-way relay communication
channel, we follow the paradigm of \emph{Wireless Physical Layer Network
Coding} (WNC) and we show that the design of novel adaptive constellations\emph{
}is desirable to exploit the promising performance of WNC in WBN.
We introduce a \emph{systematic constellation design} algorithm and
we show that the proposed constellations are capable of outperforming
conventional point-to-point modulations in WBN over the whole range
of channel SNRs. To further support the applicability of the proposed
constellations in a practical system, we integrate a simple binary
channel coding scheme, providing a robust \emph{adaptive modulation
and coding scheme} for relaying in symmetric WBNs with arbitrary channel
conditions.
\end{abstract}

\subsubsection*{Keywords:}Wireless Network Coding, Butterfly Network, Relaying, Superposition
Coding and Modulation

\pagebreak
\nomenclature{WBN}{Wireless Butterfly Network}\nomenclature{WNC}{Wireless Network
Coding}\nomenclature{HW}{Hardware}\nomenclature{SNR}{Signal to Noise
Ratio}\nomenclature{SC}{Superposition Coding}\nomenclature{HSI}{Hierarchical Side
Information}\nomenclature{DEM}{Demodulator}\nomenclature{IC}{Interference
Cancellation}\nomenclature{MA}{Multiple Access}\nomenclature{AWGN}{Additive White Gaussian
Channel}\nomenclature{ASK}{Amplitude Shift Keying}\nomenclature{QAM}{Quadrature Amplitude
Modulation}\nomenclature{SER}{Symbol Error Rate}\nomenclature{FER}{Frame Error
Rate}\nomenclature{SDR}{Software Defined Radio}\nomenclature{USRP}{Universal Software Radio
Peripherals}\nomenclature{UHD}{USRP Hardware Driver }\nomenclature{CP}{Cyclic
Prefix}\nomenclature{ODFD}{Orthogonal Frequency Division Multiplexing}\nomenclature{DC}{Direct
Current}\nomenclature{MIMO}{Multiple Input Multiple Output}\nomenclature{QPSK}{Quadrature Phase
Shift Keying}\nomenclature{UDP}{User Datagram Protocol}\nomenclature{BC}{Broadcast
Channel}\nomenclature{MAC}{Multiple Access Channel}\nomenclature{SODEM}{Soft Output Demodulator}\nomenclature{CC}{Constellation
Constrained}%\settowidth{\nomlabelwidth}{AWGN} 
\printnomenclature{} 
\pagebreak

@@

\begin{itemize}
\item $\left[\mathbf{d}_{A}^{s};\mathbf{d}_{B}^{s};f(\mathbf{d}_{A}^{b},\mathbf{d}_{B}^{b})\right]$
(from the relay signal (\ref{eq:CTUpp_relay_output}))
\item $\mathbf{d}_{i}^{b}$ (from the stored signal $z_{j}$ (\ref{eq:CTUpp_Dj_received}),
after interference cancellation (IC) of known $\mathbf{d}_{i}^{s}$)
\item $\mathbf{d}_{j}^{b}$ (from $\mathbf{d}_{i}^{b}$ and $f(\mathbf{d}_{A}^{b},\mathbf{d}_{B}^{b})$,
using a standard WNC decoding \cite{Sykora-Burr_2011-TVT})
\end{itemize}
\item $D_{j}$ merges $\mathbf{d}_{j}^{b}$ with $\mathbf{d}_{j}^{s}$ to
obtain the desired $\left[\mathbf{d}_{j}^{b};\mathbf{d}_{j}^{s}\right]$.\end{itemize}
\tabularnewline
\hline 
\end{tabular}\vskip1ex\caption{SC-based relaying scheme in uncoded WBN}.\label{tab:CTUpp_SC-relaying-scheme}
\end{table}

@@

\begin{itemize}
\item The constellations with ($N_{b}=0,\,N_{s}>0$) assume full multi-user
detection at the relay and hence they do not exploit the HSI channels
(Table~\ref{tab:CTUpp_SC-relaying-scheme}). Consequently, their
throughput does not depend on $\gamma_{{\rm {HSI}}}$. However, their
performance worsens with decreasing $\gamma_{{\rm {MAC}}}$ (compare
Figs.~\ref{fig:CTUpp_Throughput20_20},~\ref{fig:CTUpp_Throughput16_20})
and deteriorates quite dramatically with decreasing $\gamma_{{\rm {BC}}}$
(compare Figs.~\ref{fig:CTUpp_Throughput20_20},~\ref{fig:CTUpp_Throughput20_16}).
\item The constellations with ($N_{b}>0,\,N_{s}=0$) require a reliable
HSI channel and hence their throughput is zero for low $\gamma_{{\rm {HSI}}}$
and grows up to $N_{b}$ for high $\gamma_{{\rm {HSI}}}$ and sufficiently
high $\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {BC}}}$ (all Figs.~\ref{fig:CTUpp_Throughput16_20},~\ref{fig:CTUpp_Throughput17_17},~\ref{fig:CTUpp_Throughput20_20},~\ref{fig:CTUpp_Throughput20_16}).
\item The performance of all constellations with ($N_{b}>0,\,N_{s}>0$)
is influenced by all system SNRs ($\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {BC}}},\,\gamma_{{\rm {HSI}}}$).
Remarkably, for some particular values of $\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {BC}}}$
these constellations significantly outperform those with $N_{b}=0$
or $N_{s}=0$ in the medium $\gamma_{{\rm {HSI}}}$ region (Figs.~\ref{fig:CTUpp_Throughput17_17},~\ref{fig:CTUpp_Throughput20_16}).
Note that this is a direct consequence of superior $\rho_{\mathrm{HSI}}^{2}$
performance of these constellations\footnote{It can be shown that for constant $N_{b}+N_{s}$ the minimum Euclidean
distance $\rho_{\mathrm{HSI}}^{2}$ grows with decreasing $N_{b}$,
resulting in a more efficient utilization of \textquotedbl{}imperfect\textquotedbl{}
HSI channels \cite{Uricar-Sykora_2012,Uricar-Qian-Sykora-Mow_2014-IEEE-T-WC}
and consequently in a more reliable overheard HSI for the constellations
with lower (non-zero) $N_{b}$.} for constant $N_{b}+N_{s}$, i.e. for constant bits/constellation
symbol transmitted by the sources.
\begin{figure}
\begin{centering}
\hspace*{-0.04\columnwidth}\includegraphics[width=0.95\columnwidth]{fig/Throughput_HSI_XOR_MAC16_BC20_N2}\vskip-2ex
\par\end{centering}

\caption{Comparison of throughput as a function of $\gamma_{{\rm {HSI}}}$
for $\gamma_{{\rm {MAC}}}=16\,\mathrm{dB},\,\gamma_{{\rm {BC}}}=20\,\mathrm{dB}$
and given $(N_{b},N_{s})$. \label{fig:CTUpp_Throughput16_20}}
\end{figure}
\begin{figure}
\begin{centering}
\hspace*{-0.04\columnwidth}\includegraphics[width=0.95\columnwidth]{fig/Throughput_HSI_XOR_MAC17_BC17_N2}\vskip-2ex
\par\end{centering}

\caption{Comparison of throughput as a function of $\gamma_{{\rm {HSI}}}$
for $\gamma_{{\rm {MAC}}}=17\,\mathrm{dB},\,\gamma_{{\rm {BC}}}=17\,\mathrm{dB}$
and given $(N_{b},N_{s})$.\label{fig:CTUpp_Throughput17_17}}
\end{figure}
\begin{figure}
\begin{centering}
\hspace*{-0.04\columnwidth}\includegraphics[width=0.95\columnwidth]{fig/Throughput_HSI_XOR_MAC20_BC20_N2}\vskip-2ex
\par\end{centering}

\caption{Comparison of throughput as a function of $\gamma_{{\rm {HSI}}}$
for $\gamma_{{\rm {MAC}}}=20\,\mathrm{dB},\,\gamma_{{\rm {BC}}}=20\,\mathrm{dB}$
and given $(N_{b},N_{s})$. \label{fig:CTUpp_Throughput20_20}}
\end{figure}
\begin{figure}
\begin{centering}
\hspace*{-0.04\columnwidth}\includegraphics[clip,width=0.95\columnwidth]{fig/Throughput_HSI_XOR_MAC20_BC16_N2}\vskip-2ex
\par\end{centering}

@@

\pagebreak{}

@@

\section*{Funding}

This work has been performed in the framework of the European research
project DIWINE, which is partly funded by the European Union under
its FP7 ICT Objective 1.1 - The Network of the Future. Work of Pavel
Prochazka and Tomas Uricar was also partially supported by the Grant
Agency of the Czech Technical University in Prague, SGS project SGS16/094/OHK3/1T/13.

\bibliographystyle{ieeetr}
\bibliography{bib/bib,bib/uricar,bib/uricar-papers,bib/uricar-books,bib/sykora,bib/books}

\end{document}

>>>>

%% LyX 2.1.5 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[letterpaper,english,10pt]{bmc_template/bmcart}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{array}
\usepackage{float}
\usepackage{units}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[dvips]{graphicx}
\usepackage{setspace}
\usepackage{wasysym}
\usepackage{nomencl}
% the following is useful when we have the old nomencl.sty package
\providecommand{\printnomenclature}{\printglossary}
\providecommand{\makenomenclature}{\makeglossary}
\makenomenclature
\doublespacing

@@

\usepackage{babel}
\begin{document}
\begin{frontmatter}

\begin{fmbox}
\dochead{Research Article}

@@

\author[
   addressref={aff1},
   email={pavel@prochazka.info}
]{\inits{PP}\fnm{Pavel} \snm{Prochazka}}
\author[
   addressref={aff1},
   email={uricatom@fel.cvut.cz}
]{\inits{TU}\fnm{Tomas} \snm{Uricar}}
\author[
   addressref={aff2},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff2},                       % id of corresponding address, if any
   %noteref={n1},                        % id's of article notes, if any
   email={david.halls@toshiba-trel.com}   % email address
]{\inits{DH}\fnm{David} \snm{Halls}}
\author[
   addressref={aff1},
   email={jan.sykora@fel.cvut.cz}
]{\inits{JS}\fnm{Jan} \snm{Sykora}}

\address[id=aff1]{%                           % unique id
  \orgname{Czech Technical University in Prague}, % university, etc
  \street{Technicka 2},                     %
  \postcode{16627}                                % post or zip code
  \city{Prague},                              % city
  \cny{CZ}                                    % country
}
\address[id=aff2]{%
  \orgname{Toshiba -- Telecommunication Research Laboratory},
  \street{32 Queen Square},
  \postcode{BS1 4ND}
  \city{Bristol,},
  \cny{UK}
}
\end{fmbox}% comment this for two column layout

\begin{abstractbox}

\begin{abstract} % abstract
In this paper we focus on the design of a reliable communication scheme
for generic 5-node Wireless Butterfly Networks (WBN). Since the WBN
represents a natural extension of the basic 2-way relay communication
channel, we follow the paradigm of \emph{Wireless Physical Layer Network
Coding} and we show that the design of novel adaptive constellations\emph{
} is desirable to exploit the promising performance of wireless network coding in WBN.
We introduce a \emph{systematic constellation design} algorithm and
we show that the proposed constellations are capable of outperforming
conventional point-to-point modulations in WBN over the whole range
of channel signal to noise ratios. To further support the applicability of the proposed
constellations in a practical system, we integrate a simple binary
channel coding scheme, providing a robust \emph{adaptive modulation
and coding scheme} for relaying in symmetric WBNs with arbitrary channel
conditions.
\end{abstract}

\begin{keyword}
\kwd{wireless network coding}
\kwd{butterfly network}
\kwd{relaying}
\kwd{superposition coding}
\kwd{superposition modulation}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}

\end{frontmatter}

@@

\begin{itemize}
\item $\left[\mathbf{d}_{A}^{s};\mathbf{d}_{B}^{s};f(\mathbf{d}_{A}^{b},\mathbf{d}_{B}^{b})\right]$
(from the relay signal (\ref{eq:CTUpp_relay_output}))
\item $\mathbf{d}_{i}^{b}$ (from the stored signal $z_{j}$ (\ref{eq:CTUpp_Dj_received}),
after interference cancellation (IC) of known $\mathbf{d}_{i}^{s}$)
\item $\mathbf{d}_{j}^{b}$ (from $\mathbf{d}_{i}^{b}$ and $f(\mathbf{d}_{A}^{b},\mathbf{d}_{B}^{b})$,
using a standard WNC decoding \cite{Sykora-Burr_2011-TVT})
\end{itemize}
\item $D_{j}$ merges $\mathbf{d}_{j}^{b}$ with $\mathbf{d}_{j}^{s}$ to
obtain the desired $\left[\mathbf{d}_{j}^{b};\mathbf{d}_{j}^{s}\right]$.\end{itemize}
\tabularnewline
\hline 
\end{tabular}vskip1ex\caption{SC-based relaying scheme in uncoded WBN}.\label{tab:CTUpp_SC-relaying-scheme}
\end{table}

@@

\begin{itemize}
\item The constellations with ($N_{b}=0,\,N_{s}>0$) assume full multi-user
detection at the relay and hence they do not exploit the HSI channels
(Table~\ref{tab:CTUpp_SC-relaying-scheme}). Consequently, their
throughput does not depend on $\gamma_{{\rm {HSI}}}$. However, their
performance worsens with decreasing $\gamma_{{\rm {MAC}}}$ (compare
Figs.~\ref{fig:CTUpp_Throughput20_20},~\ref{fig:CTUpp_Throughput16_20})
and deteriorates quite dramatically with decreasing $\gamma_{{\rm {BC}}}$
(compare Figs.~\ref{fig:CTUpp_Throughput20_20},~\ref{fig:CTUpp_Throughput20_16}).
\item The constellations with ($N_{b}>0,\,N_{s}=0$) require a reliable
HSI channel and hence their throughput is zero for low $\gamma_{{\rm {HSI}}}$
and grows up to $N_{b}$ for high $\gamma_{{\rm {HSI}}}$ and sufficiently
high $\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {BC}}}$ (all Figs.~\ref{fig:CTUpp_Throughput16_20},~\ref{fig:CTUpp_Throughput17_17},~\ref{fig:CTUpp_Throughput20_20},~\ref{fig:CTUpp_Throughput20_16}).
\item The performance of all constellations with ($N_{b}>0,\,N_{s}>0$)
is influenced by all system SNRs ($\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {BC}}},\,\gamma_{{\rm {HSI}}}$).
Remarkably, for some particular values of $\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {BC}}}$
these constellations significantly outperform those with $N_{b}=0$
or $N_{s}=0$ in the medium $\gamma_{{\rm {HSI}}}$ region (Figs.~\ref{fig:CTUpp_Throughput17_17},~\ref{fig:CTUpp_Throughput20_16}).
Note that this is a direct consequence of superior $\rho_{\mathrm{HSI}}^{2}$
performance of these constellations\footnote{It can be shown that for constant $N_{b}+N_{s}$ the minimum Euclidean
distance $\rho_{\mathrm{HSI}}^{2}$ grows with decreasing $N_{b}$,
resulting in a more efficient utilization of \textquotedbl{}imperfect\textquotedbl{}
HSI channels \cite{Uricar-Sykora_2012,Uricar-Qian-Sykora-Mow_2014-IEEE-T-WC}
and consequently in a more reliable overheard HSI for the constellations
with lower (non-zero) $N_{b}$.} for constant $N_{b}+N_{s}$, i.e. for constant bits/constellation
symbol transmitted by the sources.
\begin{figure}
\begin{centering}
\hspace*{-0.04\columnwidth}\includegraphics[width=0.95\columnwidth]{fig/Throughput_HSI_XOR_MAC16_BC20_N2}
\par\end{centering}

\caption{Comparison of throughput as a function of $\gamma_{{\rm {HSI}}}$
for $\gamma_{{\rm {MAC}}}=16\,\mathrm{dB},\,\gamma_{{\rm {BC}}}=20\,\mathrm{dB}$
and given $(N_{b},N_{s})$. \label{fig:CTUpp_Throughput16_20}}
\end{figure}
\begin{figure}
\begin{centering}
\hspace*{-0.04\columnwidth}\includegraphics[width=0.95\columnwidth]{fig/Throughput_HSI_XOR_MAC17_BC17_N2}
\par\end{centering}

\caption{Comparison of throughput as a function of $\gamma_{{\rm {HSI}}}$
for $\gamma_{{\rm {MAC}}}=17\,\mathrm{dB},\,\gamma_{{\rm {BC}}}=17\,\mathrm{dB}$
and given $(N_{b},N_{s})$.\label{fig:CTUpp_Throughput17_17}}
\end{figure}
\begin{figure}
\begin{centering}
\hspace*{-0.04\columnwidth}\includegraphics[width=0.95\columnwidth]{fig/Throughput_HSI_XOR_MAC20_BC20_N2}
\par\end{centering}

\caption{Comparison of throughput as a function of $\gamma_{{\rm {HSI}}}$
for $\gamma_{{\rm {MAC}}}=20\,\mathrm{dB},\,\gamma_{{\rm {BC}}}=20\,\mathrm{dB}$
and given $(N_{b},N_{s})$. \label{fig:CTUpp_Throughput20_20}}
\end{figure}
\begin{figure}
\begin{centering}
\hspace*{-0.04\columnwidth}\includegraphics[clip,width=0.95\columnwidth]{fig/Throughput_HSI_XOR_MAC20_BC16_N2}
\par\end{centering}

@@

\begin{backmatter}

\nomenclature{WBN}{Wireless Butterfly Network}\nomenclature{WNC}{Wireless Network
Coding}\nomenclature{HW}{Hardware}\nomenclature{SNR}{Signal to Noise
Ratio}\nomenclature{SC}{Superposition Coding}\nomenclature{HSI}{Hierarchical Side
Information}\nomenclature{DEM}{Demodulator}\nomenclature{IC}{Interference
Cancellation}\nomenclature{MA}{Multiple Access}\nomenclature{AWGN}{Additive White Gaussian
Channel}\nomenclature{ASK}{Amplitude Shift Keying}\nomenclature{QAM}{Quadrature Amplitude
Modulation}\nomenclature{SER}{Symbol Error Rate}\nomenclature{FER}{Frame Error
Rate}\nomenclature{SDR}{Software Defined Radio}\nomenclature{USRP}{Universal Software Radio
Peripherals}\nomenclature{UHD}{USRP Hardware Driver }\nomenclature{CP}{Cyclic
Prefix}\nomenclature{ODFD}{Orthogonal Frequency Division Multiplexing}\nomenclature{DC}{Direct
Current}\nomenclature{MIMO}{Multiple Input Multiple Output}\nomenclature{QPSK}{Quadrature Phase
Shift Keying}\nomenclature{UDP}{User Datagram Protocol}\nomenclature{BC}{Broadcast
Channel}\nomenclature{MAC}{Multiple Access Channel}\nomenclature{SODEM}{Soft Output Demodulator}\nomenclature{CC}{Constellation
Constrained}%\settowidth{\nomlabelwidth}{AWGN} 
\printnomenclature{}

@@

\section*{Competing Interests}
The authors hereby confirm, having read SpringerOpen's guidance on competing interests, that we have no competing interests in the manuscript.

\section*{Funding}
This work has been performed in the framework of the European research
project DIWINE, which is partly funded by the European Union under
its FP7 ICT Objective 1.1 - The Network of the Future. Work of Pavel
Prochazka and Tomas Uricar was also partially supported by the Grant
Agency of the Czech Technical University in Prague, SGS project SGS16/094/OHK3/1T/13.

\section*{Availability of Data and Materials}

\bibliographystyle{bmc_template/bmc-mathphys}
\bibliography{bib/bib,bib/uricar,bib/uricar-papers,bib/uricar-books,bib/sykora,bib/books}

\end{backmatter}
\end{document}
